Bayesian Hierarchical Modeling: An Introduction and Reassessment

### 2022-11-13 manuscript.Rmd
Logmodel comparison

author: 
  - name          : "Myrthe Veenman"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Wassenaarseweg 52, Leiden"
    email         : "myrthe.veenman@gmail.com"
  - name          : "Angelika M. Stefan"
    affiliation   : "2"
    corresponding : no    # Define only one corresponding author
  - name          : "Julia M. Haaf"
    affiliation   : "2"
    corresponding : no    # Define only one corresponding author

affiliation:
  - id            : "1"
    institution   : "Leiden University"
  - id            : "2"
    institution   : "University of Amsterdam"

@@

In hierarchical analyses, the general or overall effect is called fixed effect, and the individual-specific deviations from this effect are called random effects (Rouder & Lu, 2005). We can also estimate individual effects, not as deviations from an overarching effect, but as effects for each individual, as obtained from a per-participant analysis. In this tutorial, we will use the terms general effect, individual deviation, and individual effects instead of fixed and random effects, because we believe these terms are better suited for the interpretation of our analyses.

@@

Finally, Figure \@ref(fig:modelestimatesfigure3) shows how the relationship between individuals' digit effects.^[The design and code for this figure are by @j_van_langen_2020_3715576.] The posterior means of individual effects are represented by the dots. The right side of each plot represents the distribution of the individual estimates. Note that this distribution is not the posterior distribution of the general effect. It rather corresponds to the left side of Figure 6. The two panels of Figure \@ref(fig:modelestimatesfigure3) show that the estimates are almost identical between `Stan` and `brms`. This means that differences in the parametrization between the packages did not influence the results of the parameter estimation.

@@

\begin{equation}
\text{BF}_{\text{nl}} = \frac{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{n}, \bm{\theta_{\text{n}}})f(\bm{\theta_{\text{n}}})d\bm{\theta_{\text{n}}}}{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{l}, \bm{\theta_{\text{l}}})f(\bm{\theta_{\text{l}}})d\bm{\theta_{\text{l}}}}. 
\end{equation}

@@

This tutorial showed the application of two R packages: `rstan` and `brms`. Which package to use depends mostly on the complexity of the model and your technical background. For the non-technical user, we recommend using `brms` because of its intuitive model specification. However, the default prior setup can be unsuitable for a specific research context and might need to be adapted. Additionally, users are limited to predefined parametrizations. `rstan` offers the most flexibility in model and prior specification. This flexibility comes with a cost: if the prior or the model is not well defined, results can be misleading.

>>>>

author: 
  - name          : "Myrthe Veenman"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Wassenaarseweg 52, Leiden"
    email         : "myrthe.veenman@gmail.com"
  - name          : "Angelika M. Stefan"
    affiliation   : "2"
    corresponding : no    # Define only one corresponding author
  - name          : "Julia M. Haaf"
    affiliation   : "3"
    corresponding : no    # Define only one corresponding author

affiliation:
  - id            : "1"
    institution   : "Leiden University"
  - id            : "2" 
    institution   : "Universität der Bundeswehr München"
  - id            : "3"
    institution   : "University of Amsterdam"

@@

\textcolor{blue}{In hierarchical analyses, the general or overall effect is sometimes called fixed effect, and the individual-specific deviations from this effect are in that case referred to as random effects (Rouder \& Lu, 2005). We can also estimate individual effects, not as deviations from an overarching effect, but as effects for each individual, as obtained from a per-participant analysis. Therefore, more common terms that suit the interpretation of the analyses better are general effect, individual deviation, and individual effects instead of fixed and random effects. We will use the terms general effect, individual deviation, and individual effect throughout this tutorial.}

@@

Finally, Figure \@ref(fig:modelestimatesfigure3) shows how the relationship between individuals' digit effects.^[The design and code for this figure are by @j_van_langen_2020_3715576.] The posterior means of individual effects are represented by the dots. The right side of each plot represents the distribution of the individual estimates. Note that this distribution is not the posterior distribution of the general effect. It rather corresponds to the left side of Figure 6.

@@

\begin{equation}
\text{BF}_{\text{nl}} = \frac{P(\bm{Y} | \mathcal{M}_n)}{P(\bm{Y} | \mathcal{M}_{\text{l}}) = \frac{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{n}, \bm{\theta_{\text{n}}})f(\bm{\theta_{\text{n}}})d\bm{\theta_{\text{n}}}}{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{l}, \bm{\theta_{\text{l}}})f(\bm{\theta_{\text{l}}})d\bm{\theta_{\text{l}}}}. 
\end{equation}

@@

This tutorial showed the application of two R packages: `rstan` and `brms`. Which package to use depends mostly on the complexity of the model and your technical background. For the non-technical user, we recommend using `brms` because of its intuitive model specification. However, the default prior setup can be unsuitable for a specific research context and (virtually) always needs to be adapted. Additionally, users are limited to predefined parametrizations. `rstan` offers the most flexibility in model and prior specification. This flexibility comes with a cost: if the prior or the model is not well defined, results can be misleading.

### 2022-11-13 manuscript.tex
Logmodel comparison

In hierarchical analyses, the general or overall effect is called fixed effect, and the individual-specific deviations from this effect are called random effects (Rouder \& Lu, 2005). We can also estimate individual effects, not as deviations from an overarching effect, but as effects for each individual, as obtained from a per-participant analysis. In this tutorial, we will use the terms general effect, individual deviation, and individual effects instead of fixed and random effects, because we believe these terms are better suited for the interpretation of our analyses.

@@

Finally, Figure \ref{fig:modelestimatesfigure3} shows how the relationship between individuals' digit effects.\footnote{The design and code for this figure are by Langen (2020).} The posterior means of individual effects are represented by the dots. The right side of each plot represents the distribution of the individual estimates. Note that this distribution is not the posterior distribution of the general effect. It rather corresponds to the left side of Figure 6. The two panels of Figure \ref{fig:modelestimatesfigure3} show that the estimates are almost identical between \texttt{Stan} and \texttt{brms}. This means that differences in the parametrization between the packages did not influence the results of the parameter estimation.

@@

This tutorial showed the application of two R packages: \texttt{rstan} and \texttt{brms}. Which package to use depends mostly on the complexity of the model and your technical background. For the non-technical user, we recommend using \texttt{brms} because of its intuitive model specification. However, the default prior setup can be unsuitable for a specific research context and might need to be adapted. Additionally, users are limited to predefined parametrizations. \texttt{rstan} offers the most flexibility in model and prior specification. This flexibility comes with a cost: if the prior or the model is not well defined, results can be misleading.

>>>>

\textcolor{blue}{In hierarchical analyses, the general or overall effect is sometimes called fixed effect, and the individual-specific deviations from this effect are in that case referred to random effects (Rouder \& Lu, 2005). We can also estimate individual effects, not as deviations from an overarching effect, but as effects for each individual, as obtained from a per-participant analysis. Therefore, more common terms that suit the interpretation of the analyses better are general effect, individual deviation, and individual effects instead of fixed and random effects. We will use the terms general effect, individual deviation, and individual effect throughout this tutorial.}

@@

Finally, Figure \ref{fig:modelestimatesfigure3} shows how the relationship between individuals' digit effects.\footnote{The design and code for this figure are by Langen (2020).} The posterior means of individual effects are represented by the dots. The right side of each plot represents the distribution of the individual estimates. Note that this distribution is not the posterior distribution of the general effect. It rather corresponds to the left side of Figure 6.

@@

This tutorial showed the application of two R packages: \texttt{rstan} and \texttt{brms}. Which package to use depends mostly on the complexity of the model and your technical background. For the non-technical user, we recommend using \texttt{brms} because of its intuitive model specification. However, the default prior setup can be unsuitable for a specific research context and (virtually) always needs to be adapted. Additionally, users are limited to predefined parametrizations. \texttt{rstan} offers the most flexibility in model and prior specification. This flexibility comes with a cost: if the prior or the model is not well defined, results can be misleading.
